---
title: "WeightK"
output: html_document
---
libraries (Remember to instal all the packages)
```{r eval=FALSE}
library(MixSim) #To create the cluster data
library(NbClust) #30 stopping rules 
library(plyr) #For data editing
library(tidyr) #For data editing

```


```{r}
#Set WD to data here 
setwd("C:/Users/Camil/Desktop/Internship/Code/Data")

#Save data as variable d. For NbClust each dimension has a column 
d <- read.csv("ResultSimK.csv") 

#Load weights
weight <- read.csv("Weights.csv") #Data from 1000 simulations of clusters, where NbClust has calculated K from 30 stopping rules
```
Run NbClust on the data:
Depending on how big the data is, this make take a long time (maybe hours). If you don't have time for that, change index from alllong to all and delete GAP, Gamma, Gplus and Tau from the weihts dataset. 
```{r}
NbEstimations <-NbClust(d, distance = "euclidean", min.nc = 2, max.nc = 10, method = "ward.D", index = "alllong") 

Estomations=NbEstimations$Best.nc[1,]
```

Here I create a loop which will subset the estimations into groups based on the estimation. So all stopping rules that estimated 3 cluseters in one subset and all stopping rules that guseed for 4 clustsers in a different subset. For each subset I sum the weights assigned to each stopping rule. The number of clusters with the highest number is the final estimation for the amount of clusters in the dataset. The numbers are abitrary. 
```{r}
SummedGuesses = data.frame(0) #An empty dataframe to be filled in the loop
CurrentGuess = data.frame(0) #An empty dataframe to be filled in the loop
Score = data.frame(0) #An empty dataframe to be filled in the loop



data = data.frame(Guess=cbind("Estimations" = Estomations, "weight"=weight$weight.score.100)) #Estimations from nbclust and weights in the same dataframe. 

Guesses = data[,1] %>% #Find all unique guesses 
  unique() %>%
  na.omit() %>% # Remove NAs 
  data.frame() #Make into dataframe

#Sum all the weights for each unique guess 
for (i in 1:nrow(Guesses)){
  
  CurrentGuess <- data.frame(subset(data, data$Guess.Estimations == Guesses[i,])) #Subset into rules that guessed the same number
  Score <- sum(as.numeric(CurrentGuess[,2])) # sum the weights in the subset 
  SummedGuesses <- cbind(SummedGuesses, Score)  #Make a dataframe that collects all the summed results 
  
  
}

SummedGuesses = t(SummedGuesses) #I transpose the data

Guess = data.frame(cbind(Guess=Guesses, Score=SummedGuesses[-1,])) #Dataframe with the unique guesses and the summed results. The higest is the suggested number of clusters. 


plot(Guess) #Very simple plot to show the results. 
  
```

